Splitting of the homology of the punctured mapping class group

### 2019-02-18 Actiongg.tex
Implemented some corrections by cfb, created gitignore file

\section{Action of $\gg$}
We now turn back to \emph{homology} of $\cms$; we will first prove the full statement of theorem
\ref{thm:Hbms*as*ggrep} in bidigrees $(*,m)$ with $*=m$, and then extend it to other bidegrees.

\subsection{Bidigrees of the form $(m,m)$}
In this subsection we construct a $\gg-$~equivariant isomorphism 
\[
\psi_m\colon \Sym_m(\H))\to H_m(\cms).
\]
% the former group is isomorphic to $\Sym_m(H^1(\S))$ and the second to $H^m(\cms)$,
% both by Poincaré-Lefschetz duality.

We already know that these $\Z_2-$vector spaces have the same dimension:
indeed $\Sym_m(\H)$ is precisely the summand in bidegree $(m,m)$ in equation
\ref{eq:isovectorspaces}, using that a monomial
 whose
weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
$Q^i\epsilon$.
% i.e. it must lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
% is $m$ it must lie in $\Sym_m(\H)$.
% 
% The rough idea in constructing the map $\psi_m$ is the following: $\Sym_m(\H)$ is generated
% by elements of the form $[c_1]\cdot\ldots\cdot[c_m]$, where $c_1,\dots,c_m\colon\Sone\to\mrS$ are simple
% closed curves; we consider the product map 
% \[
% c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to \SP^m(\mrS)
% \]
% which carries the fundamental class of the $m-$fold torus to some class in
% $H_m(\SP^m(\mrS))$; we then lifts this class, in a nice way, to a class
% in $H_m(\cms)$.

@@

\begin{defn}
\label{defn:cCm}
Let $\cC^m$ be the the (discrete) set of isotopy classes of $m-$tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
such that
%\begin{itemize}
%  \item each $c_r\colon\Sone\to\mrS$ is smooth and transverse to each $\U_i$ and $\V_i$;
% \item
 any two curves $c_i,c_j$ intersect each other only inside $\D$.

% \item each curve $c_i$ is transverse to $\partial\D$.
%\end{itemize}
Curves are seen as maps $\Sone\to\mrS$, and the intersection of two curves is the intersection of their
images.

Two $m-$tuples of curves are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
transforming one $m-$tuple into the other. In particular $\cC^m$ is more than countable.

@@

We denote by $\ZcC{m}$ the free $\Z_2-$vector space with basis $\cC^m$.
There is a canonical surjective map $\pr_m\colon\ZcC{m}\to\Sym_m(\H)$ given by
\[
 \pr_m(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m],
\]
where $[c_i]\in\H$ is the fundamental class of the curve $c_i$.

The group $\gg$ acts both on $\ZcC{m}$, by acting on its basis $\cC^m$, and
on $\Sym_m(\H)$, symplectically. The map $\pr_m$ is $\gg-$equivariant.
\end{defn}

\begin{figure}\centering
 \includegraphics[scale=0.7]{figures/defcCm.png}
 \caption{A multicurve in $\cC^5$.}
\label{fig:defcCm}
\end{figure}

% Our aim is to define a $\gg-$equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
% factors through $\pr$ to give the desired $\gg-$equivariant map $\psi$; we will then show that
% $\psi$ is an isomorphism.

% The map $\tpsi_m$ that we construct will have two additional properties. The first
% is the following: for any
% $(c_1,\dots,c_m)$ representing a class in $\cC^m$, the homology
% class $\tpsi_m(c_1,\dots,c_m)$ can be represented by a singular cycle in $\cms$ which
% is \emph{supported} on the union $\D\cup c_1\cup \dots\cup c_m$, meaning that
% this singular cycle only hits configurations in $\cms$ whose points actually
% lie in $\D\cup c_1\cup \dots\cup c_m$.
% 
% For the second property we need a definition
\begin{defn}
 \label{defn:cmsD}
 The space $\cmsD$ is the subspace of $\SP^{m}(\mrS)$ of configurations where all points
 having multiplicity $\geq 2$ lie inside $D$.

 Recall that the $m-$fold symmetric product of an open surface is a $2m-$dimensional manifold;
 the space $\cmsD$ is an open subset
 of $\SP^m(\mrS)$, hence $\cmsD$ is also a manifold.
 See figure \ref{fig:defcmsD}.

 There is a natural, open inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
  $\j_m\colon \ZcC{m}\to H_m(\cmsD)$ defined as follows: for a class $(c_1,\dots,c_m)\in\cC^m$
  the composition
  \[
   \begin{CD}
    \pa{\Sone}^{\times m} @>c_1\times\dots\times c_m >> (\mrS)^{\times m} @>>> \SP^m(\mrS)
   \end{CD}
  \]
has image in the subspace $\cmsD$; we define $\j_m(c_1,\dots,c_m)$ as the image
of the fundamental class
of the $m-$fold torus in $H_m(\cmsD)$. The result doesn't change
if we substitute $(c_1,\dots,c_m)$ with another $m-$tuple in the same multicurve.

We call $[c_1]\cdot\ldots\cdot[c_m]$ the image in $\cmsD$ of the fundamental cycle of the
$m-$fold torus: this cycle represents the class $\j_m(c_1,\dots,c_m)$ and it is \emph{supported}
on the union $c_1\cup\dots\cup c_m$, meaning that it hits configurations in $\cmsD$ of points of $\mrS$
lying in this union.

The group $\Diff(\S,\D\cup\partial\S)$ acts on $\cmsD$, and there is an induced action of $\gg$
on $H_m(\cmsD)$. The map $\j_m$ is $\gg-$equivariant.
\end{defn}

\begin{lem}
 \label{lem:cms->cmsDinj}
 The inclusion $\iota_m\colon\cms\to\cmsD$ induces an injective map
 \[
  \pa{\iota_m}_*\colon H_m(\cms)\to H_m(\cmsD).
 \]
\end{lem}
\begin{proof}
 It is equivalent to prove that the map $\iota_m^*\colon H^m(\cmsD)\to H^m(\cms)$ is surjective,
 or, using Poincaré-Lefschetz duality, that the map $\tH_m(\cmsD^c)\to \tH_m(\cms^c)$
 is surjective: this last map is induced by the map $\cmsD^c\to\cms^c$ collapsing the subspace
 $\cmsD^c\setminus\cms$ to $\infty$.

 Recall from lemma \ref{lem:gensymchain} that a $\Z_2-$basis for $\tH_m(\cms^c)$
 is given by classes $[p,(\alpha_j),(u_i,v_i)]$ with $p=0$, $\alpha_j=0$ for all $j$,
 and $\sum_{i=1}^g (u_i+v_i)=m$. Here and in the following we abbreviate $[0,(u_i,v_i)]$ for
 such classes.

 The class $[0,(u_i,v_i)]$ is represented by a generalised symmetric chain consisting of only one cell,
 the cell $e^{\tup}$
 with $\tup=(0,(u_i,v_i))$.

 In particular there is a map of pairs
 $\phi^{\tup}\colon(\Delta^{\tup},\partial\Delta^{\tup})\to (\cms^c,\infty)$, and $[0,(u_i,v_i)]$
 is the image along this map of the fundamental class of $H_m\pa{\Delta^{\tup},\partial\Delta^{\tup}}$.

 It is straightforward to check that the map $\phi^{\tup}$ factors through
 $(\cmsD,\infty)$, as a map of pairs; surjectivity of $\tH_m(\cmsD^c)\to \tH_m(\cms^c)$ follows.
\end{proof}

We have the following diagram of $\gg$-equivariant maps, where the full arrows are those
that we have already constructed, and we still have to prove the existence of the dashed arrows
\begin{equation}
\label{eq:fulldasheddiagram}
\begin{tikzcd}[column sep=8em,row sep=5em]
  \ZcC{m} \ar[r,"\pr_m",two heads]\ar[d,dashed, swap, "\tpsi_m"]\ar[dr,"\j_m",near start]
  & \Sym_m(\H)\ar[d,dashed ]\ar[dl,dashed,very near start,"\psi_m"]\\
  H_m(\cms)\ar[r,swap,"\pa{\iota_m}*"hook] & H_m\pa{\cmsD}
 \end{tikzcd}
\end{equation}

We now prove that the map $\j_m$ lifts along $(\iota_m)_*$
to a $\gg-$equivariant map $\tpsi_m$ as in the diagram.
Since $(\iota_m)_*$ is injective by lemma \ref{lem:cms->cmsDinj}, it suffices to prove that $\j_m$
lands in the image of $(\iota_m)_*$, and this last statement doesn't depend on how $\gg$ acts on these groups.

@@

\begin{lem}
 \label{lem:tpsiwithproperties}
For each $m-$tuple of curves $(c_1,\dots,c_m)$ representing a class in $\cC^m$
and for each open neighborhood $\N\subset\mrS$ of
$\D\cup c_1\cup\dots\cup c_m$, there is a singular cycle $\fc=\tpsi_m(c_1,\dots,c_m)$
in $\cms$ with the following properties:
\begin{itemize}
 \item the cycle $\fc$ is \emph{supported} on $\N$, i.e.,
this singular cycle only hits configurations of $m$ distinct points of $\mrS$ that actually lie in $\N$;
\item $\pa{\iota_m}_*(\fc)$ represents the homology class $\j_m(c_1,\dots,c_m)\in H_m\pa{\cmsD}$;
\item the two cycles $\pa{\iota_m}_*(\fc)$ and $[c_1]\cdot\ldots\cdot[c_m]$
are connected by a homology in $\cmsD$ which is supported on $\N$ (the word
\emph{homology} denotes here a $(m+1)-$singular chain whose boundary is the difference between the two cycles).
\end{itemize}
\end{lem}

For $m=0$ both $\ZcC{0}$ and $H_0(C_0(\S))$ are
isomorphic to $\Z_2$ and there is nothing to show; for $m=1$
we have a canonical identification $H_1(C_1(\S))\simeq\H\simeq\Sym_1(\H)$, so we take $\tpsi_1=\pr_1$;
obviously for all $c_1$ representing a class in $\cC^1$, the homology class $\pr_1(c_1)\in\H$
is represented by a cycle supported on $c_1$,
and in this case the cycles $\iota_*\pa{\tpsi(c_1)}$ and $\pr_1(c_1)=\j_m(c_1)$ coincide.

Let now $m\geq 1$ and in the following fix a class $(c_1,\dots, c_{m+1})\in\cC^m$.
% and suppose that we have proved the existence of the lift $\tpsi_m\colon \ZcC{m}\to H_m(\cms)$
% with the aforementioned property.

\begin{defn}
\label{defn:variationsCm}
We introduce a few variations of the notion of configuration space; see figure \ref{fig:defcmsD}.
\begin{itemize} 
 \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots,P_m}}$ with $\bar P\neq P_i$ for all $i$; in other words it is
 the space of configurations of $m+1$ points, one of which is \emph{white} (meaning that
 it is distinguishable from the other points), whereas the other are \emph{black} and not distinguishable
 from each other.
%  \item The space $\fmsD$ is the subspace of $\mrS\times\cmsD$ containing all configurations
%  $(q;\set{p_1,\dots,p_m})$ where either $q\in\D$ may coincide with some $p_i$ (which can appear
%  with multipliciy $\geq 1$), or $q\not\in\D$ must be distinct from all $p_i$'s.
%  Again $q$ is thewhite point
 \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots, P_m}}$ where either $\bar P\in\D$ may coincide with \emph{exactly}
 one $P_i$, or
 $\bar P\not\in \D$ must be distinct from all $P_i$'s.
 Again $\bar P$ is called the white point.
 %We see that $C_{1,m}\subset\fmstwoD$ is an open subspace.
 \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}(\mrS)$ of configurations where either all $m+1$ points
 are distinct, or there is exactly one point \emph{inside $\D$} with multiplicity $2$ and $m-1$ other points,
 somewhere in $\mrS$, with multiplicity $1$.
\end{itemize}
 \end{defn}

\begin{figure}\centering
 \includegraphics{figures/defcmsD.png}
 \caption{A configuration in each of the space introduced in definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
 Whenever a multiplicity is not specified, it is equal to 1.}
\label{fig:defcmsD}
\end{figure}

We have the following natural inclusions:
\[
C_{1,m}(\S)\subset\fmstwoD\subset\fmsD\subset\mrS\times\cmsD\subset\mrS\times\SP^m(\mrS);
\]
\[
C_{m+1}(\S)\subset\cmstwoD\subset C_{m+1}^{\D}(\S)\subset\SP^{m+1}(\mrS).
\]
All these spaces are
manifolds of dimension $2m+2$ and all inclusions are open.

@@

We can multiply both
cycles and the homology between them by the cycle $[c_1]$: the result are the two homologous cycles
$[c_1]\cdot \fc$ and $[c_1]\cdot\ldots\cdot[c_{m+1}]$ in $C_{m+1}^{\D}(\S)$: both cycles and the
homology between them are supported on $\N$. Notice now that the
cycle $[c_1]\cdot\fc$ lives in $\cmstwoD$, so the first lift is done and we can now
deal with the second lift.

There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the white point
into a black point. This map restricts to %maps $\fmstwoD\to\cmstwoD$ and 
a map $C_{1,m}(\S)\to C_{m+1}(\S)$, so that
% the latter being even a $m+1-$fold covering map;
we have a commutative diagram
 \begin{equation}\label{eq:cmstwodiagram}
  \begin{CD}
   C_{1,m}(\S) @>\subset >> \fmstwoD %@>\subset >> \fmsD
\\   @V\p VV @V\p VV %@V\p VV
\\   C_{m+1}(\S) @>\subset >> \cmstwoD %@>\subset >> C_{m+1}^D(\S)
   \end{CD}
\end{equation}

\begin{defn}
\label{defn:falsediagonals} 
Let 
\[
\Dmone=\cmstwoD\setminus C_{m+1}(\S)
\]
and similarly
\[
\Donem=\fmstwoD\setminus C_{1,m}(\S).
\]
We notice that $\p$ restricts to a homeomorphism $\Donem\to\Dmone$. Moreover both $\Dmone\subset\cmstwoD$
and $\Donem\subset\fmstwoD$ are closed submanifolds of codimension $2$, and the map $\p$ restricts to
a $2-$fold ramified covering between their respective normal bundles.
\end{defn}

Diagram \ref{eq:cmstwodiagram} induces a commutative diagram in homology
\begin{equation}
 \label{eq:fivediagram}
\minCDarrowwidth15pt
 \begin{CD}
  @. H_{m+1}\pa{\fmstwoD} @>>> H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)}\\
  @. @V\p_*VV @V\p_*VV\\
  H_{m+1}\pa{ C_{m+1}(\S)} @>>> H_{m+1}\pa{\cmstwoD} @>>> H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}
 \end{CD}
\end{equation}

@@

We first notice that there is a lift $[c_1]\otimes\fc$ to a cycle in $\pa{\fmstwoD}$,
which is defined by declaring the point in $[c_1]\cdot\fc$ that spins around $c_1$
to be white. We then notice that the right vertical map
\[
\p_*\colon H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)} \to H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}
\]
can be rewritten, after using excision to tubular neighborhoods of $\Donem$ and $\Dmone$ respectively,
and the Thom isomorphism, as a map
\[
 H_{m-1}(\Donem)\to H_{m-1}(\Dmone).
\]
The latter map is multiplication by $2$, after identifying $\Dmone$ and $\Donem$ along $p$:
indeed the normal bundle of $\Donem$ is a
double covering of the normal bundle of $\Dmone$, hence the Thom class of the first disc
bundle corresponds to twice the Thom class of the second disc bundle. We are working
with coefficients in $\Z_2$, so multiplication by $2$ is the zero map.

Therefore the image of the cycle $[c_1]\otimes \fc$ along the diagonal of the square
in diagram \ref{eq:fivediagram} is zero; hence the image of $[c_1]\cdot\fc$ in $H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}$
is zero, hence the homology class of $[c_1]\cdot\fc$ comes from $H_{m+1}(C_{m+1}(\S))$: more
precisely there exists a cycle $\fc'$ in $C_{m+1}(\S)$ such that $\pa{\iota_{m+1}}_*(\fc')$ is homologous
to $[c_1]\cdot\fc$.

To prove lemma \ref{lem:tpsiwithproperties} we need to find a good cycle and
a good homology, namely two that are supported on $\N$: a priori both $\fc'$ and the homology
between $\pa{\iota_{m+1}}_*(\fc')$ and $[c_1]\cdot\fc$ are only supported on $\mrS$.

>>>>

\section{Action of $\gg$}
\label{sec:Actiongg}
We now turn back to \emph{homology} of $\cms$; we will first prove the full statement of Theorem
\ref{thm:Hbms*as*ggrep} in bigradings $(*,m)$ with $*=m$, and then extend it to other bigradings.

\subsection{Bigradings of the form $(m,m)$}
In this subsection we construct a $\gg$-~equivariant isomorphism 
\[
\psi_m\colon \Sym_m(\H))\to H_m(\cms).
\]
% the former group is isomorphic to $\Sym_m(H^1(\S))$ and the second to $H^m(\cms)$,
% both by Poincaré-Lefschetz duality.

We already know that these $\Z_2$-vector spaces have the same dimension:
indeed $\Sym_m(\H)$ is precisely the summand in bigrading $(m,m)$ in equation
\eqref{eq:isovectorspaces}, using that a monomial
 whose
weight is \emph{equal} and not bigger than its degree cannot contain factors of the form
$Q^i\epsilon$.
% i.e. it must lie in the factor $\Sym_{\bullet}(\H)$, and if the degree
% is $m$ it must lie in $\Sym_m(\H)$.
% 
% The rough idea in constructing the map $\psi_m$ is the following: $\Sym_m(\H)$ is generated
% by elements of the form $[c_1]\cdot\ldots\cdot[c_m]$, where $c_1,\dots,c_m\colon\Sone\to\mrS$ are simple
% closed curves; we consider the product map 
% \[
% c_1\times\dots\times c_m\colon \pa{\Sone}^{\times m}\to \SP^m(\mrS)
% \]
% which carries the fundamental class of the $m$-fold torus to some class in
% $H_m(\SP^m(\mrS))$; we then lifts this class, in a nice way, to a class
% in $H_m(\cms)$.

@@

\begin{defn}
\label{defn:DinTS} 
Recall Definition \ref{defn:TS}. We denote by $\D\subset\T(\S)\cong\mrS$ the open square $]1/4;3/4[\times]1/2,1[$.
Note that $\D$ is an open disc in $\mrS$ near $\partial\S$ and 
it is disjoint from all $\overline{\U}_i$'s and $\overline{\V}_i$'s. The interior of the surface $\S'$
is then identified with $\T(\S)\setminus [1/4,3/4]\times[1/2,1[$.
\end{defn}

\begin{defn}
\label{defn:cCm}
Let $\cC^m$ be the the (discrete) set of isotopy classes of $m$-tuples of simple closed curves $c_1,\dots,c_m\subset\mrS$
such that
%\begin{itemize}
%  \item each $c_r\colon\Sone\to\mrS$ is smooth and transverse to each $\U_i$ and $\V_i$;
% \item
 any two curves $c_i,c_j$ intersect each other only inside $\D$.

% \item each curve $c_i$ is transverse to $\partial\D$.
%\end{itemize}
Curves are seen as maps $\Sone\to\mrS$, and the intersection of two curves is the intersection of their
images. Here and in the following $\Sone$ is the unit circle in $\C$.

Two $m$-tuples of curves are isotopic if there is an ambient isotopy of $\S$ relative to $\partial\S\cup\D$
transforming one $m$-tuple into the other. In particular $\cC^m$ is more than countable.

@@

We denote by $\ZcC{m}$ the free $\Z_2$-vector space with basis $\cC^m$.
There is a canonical surjective map $\pr_m\colon\ZcC{m}\to\Sym_m(\H)$ given by
\[
 \pr_m(c_1,\dots,c_m)=[c_1]\cdot\ldots\cdot[c_m],
\]
where $[c_i]\in\H$ is the fundamental class of the curve $c_i$.

The group $\gg$ acts both on $\ZcC{m}$, by acting on its basis $\cC^m$, and
on $\Sym_m(\H)$, symplectically. The map $\pr_m$ is $\gg$-equivariant.
\end{defn}

\begin{figure}\centering
 \includegraphics[scale=0.7]{figures/defcCm.pdf}
 \caption{A multicurve in $\cC^5$.}
\label{fig:defcCm}
\end{figure}

% Our aim is to define a $\gg$-equivariant map $\tpsi\colon\ZcC{m}\to H_m(\cms)$ and show that it
% factors through $\pr$ to give the desired $\gg$-equivariant map $\psi$; we will then show that
% $\psi$ is an isomorphism.

% The map $\tpsi_m$ that we construct will have two additional properties. The first
% is the following: for any
% $(c_1,\dots,c_m)$ representing a class in $\cC^m$, the homology
% class $\tpsi_m(c_1,\dots,c_m)$ can be represented by a singular cycle in $\cms$ which
% is \emph{supported} on the union $\D\cup c_1\cup \dots\cup c_m$, meaning that
% this singular cycle only hits configurations in $\cms$ whose points actually
% lie in $\D\cup c_1\cup \dots\cup c_m$.
% 
% For the second property we need a definition
\begin{defn}
 \label{defn:cmsD}
 Recall Definition \ref{defn:symprod}.
 The space $\cmsD$ is the subspace of $\SP^m(\mrS)$ of configurations where all points
 having multiplicity $\geq 2$ lie inside $D$.
 See Figure \ref{fig:defcmsD}.

  Note that $\cmsD$ is open in $\SP^m(\mrS)$. There is an open inclusion $\iota_m\colon\cms\subset\cmsD$ and there is a natural map
  \[
  \j_m\colon \ZcC{m}\to H_m(\cmsD)
  \]
  defined as follows: for a class $(c_1,\dots,c_m)\in\cC^m$
  the composition
  \[
   \begin{CD}
    \pa{\Sone}^{\times m} @>c_1\times\dots\times c_m >> (\mrS)^{\times m} @>>> \SP^m(\mrS)
   \end{CD}
  \]
has image in the subspace $\cmsD$; we define $\j_m(c_1,\dots,c_m)$ as the image
of the fundamental class
of the $m$-fold torus in $H_m(\cmsD)$. The result does not change
if we substitute $(c_1,\dots,c_m)$ with another isotopic $m$-tuple of curvesin the same multicurve.

We call $[c_1]\cdot\ldots\cdot[c_m]$ the image in the singular chain complex of $\cmsD$ of
the fundamental cycle of the
$m$-fold torus: this cycle represents the class $\j_m(c_1,\dots,c_m)$ and it is \emph{supported}
on the union $c_1\cup\dots\cup c_m$, meaning that it hits configurations in $\cmsD$ of points of $\mrS$
lying in this union.

The group $\Diff(\S,\D\cup\partial\S)$ acts on $\cmsD$, and there is an induced action of $\gg$
on $H_m(\cmsD)$. The map $\j_m$ is $\gg$-equivariant.
\end{defn}

\begin{lem}
 \label{lem:cms->cmsDinj}
 The inclusion $\iota_m\colon\cms\to\cmsD$ induces an injective map
 \[
  \pa{\iota_m}_*\colon H_m(\cms)\to H_m(\cmsD).
 \]
\end{lem}
\begin{proof}
 It is equivalent to prove that the map $\iota_m^*\colon H^m(\cmsD)\to H^m(\cms)$ is surjective,
 or, using Poincaré-Lefschetz duality, that the map $\tH_m(\cmsD^{\infty})\to \tH_m(\cms^{\infty})$
 is surjective: this last map is induced by the map $\cmsD^{\infty}\to\cms^{\infty}$ collapsing the subspace
 $\cmsD^{\infty}\setminus\cms$ to $\infty$.

 Recall Lemma \ref{lem:gensymchain} and Definition \ref{defn:gensymchain}. A basis for $\tH_m(\cms^{\infty})$
 is given by classes $[\kappa(0,\uu,\uv)]$, with
 $\uu=(u_1,\dots, u_g)$, $\uv=(v_1,\dots,v_g)$ and
 $\sum_{i=1}^g (u_i+v_i)=m$.

 The class $[\kappa(\uu,\uv)]$ is represented by a generalised symmetric chain consisting of only one tuple $\tup=(0,\uu,\uv)$.

 In particular there is a map of pairs
 $\phi^{\tup}\colon(\Delta^{\tup},\partial\Delta^{\tup})\to (\cms^{\infty},\infty)$, and the class $[\kappa(\uu,\uv)]$
 is the image along this map of the fundamental class of $H_m\pa{\Delta^{\tup},\partial\Delta^{\tup}}$.

 It is straightforward to check that the map $\phi^{\tup}$ factors through
 $(\cmsD,\infty)$, as a map of pairs; surjectivity of $\tH_m(\cmsD^{\infty})\to \tH_m(\cms^{\infty})$ follows.
\end{proof}

We have the following diagram of $\gg$-equivariant maps, where the full arrows are those
that we have already constructed, and we still have to prove the existence of the dashed arrows
\begin{equation}
\label{eq:fulldasheddiagram}
\begin{tikzcd}[column sep=8em,row sep=5em]
  \ZcC{m} \ar[r,"\pr_m",two heads]\ar[d,dashed, swap, "\tpsi_m"]\ar[dr,"\j_m",near start]
  & \Sym_m(\H)\ar[d,dashed ]\ar[dl,dashed,very near start,"\psi_m"]\\
  H_m(\cms)\ar[r,swap,"\pa{\iota_m}*",hook] & H_m\pa{\cmsD}.
 \end{tikzcd}
\end{equation}

We now prove that the map $\j_m$ lifts along $(\iota_m)_*$
to a $\gg$-equivariant map $\tpsi_m$ as in the diagram.
Since $(\iota_m)_*$ is injective by Lemma \ref{lem:cms->cmsDinj}, it suffices to prove that $\j_m$
lands in the image of $(\iota_m)_*$, and this last statement does not depend on how $\gg$ acts on these groups.

@@

\begin{lem}
 \label{lem:tpsiwithproperties}
For each $m$-tuple of curves $(c_1,\dots,c_m)$ representing a class in $\cC^m$
and for each open neighborhood $\N\subset\mrS$ of
$\D\cup c_1\cup\dots\cup c_m$, there is a singular cycle $\fc=\tpsi_m(c_1,\dots,c_m)$
in $\cms$ with the following properties:
\begin{itemize}
 \item the cycle $\fc$ is \emph{supported} on $\N$, i.e.,
this singular cycle only hits configurations of $m$ distinct points of $\mrS$ that actually lie in $\N$;
\item $\pa{\iota_m}_*(\fc)$ represents the homology class $\j_m(c_1,\dots,c_m)\in H_m\pa{\cmsD}$;
\item the two cycles $\pa{\iota_m}_*(\fc)$ and $[c_1]\cdot\ldots\cdot[c_m]$
are connected by a homology in $\cmsD$ which is supported on $\N$ (the word
\emph{homology} denotes here a $(m+1)$-singular chain whose boundary is the difference between the two cycles).
\end{itemize}
\end{lem}

For $m=0$ both $\ZcC{0}$ and $H_0(C_0(\S))$ are
isomorphic to $\Z_2$ and there is nothing to show. For $m=1$
we have a canonical identification $H_1(C_1(\S))\simeq\H\simeq\Sym_1(\H)$, so we take $\tpsi_1=\pr_1$;
obviously for all $c_1$ representing a class in $\cC^1$, the homology class $\pr_1(c_1)\in\H$
is represented by a cycle supported on $c_1$,
and in this case the cycles $\iota_*\pa{\tpsi(c_1)}$ and $\pr_1(c_1)=\j_m(c_1)$ coincide.

Let now $m\geq 1$ and in the following fix a class $(c_1,\dots, c_{m+1})\in\cC^{m+1}$.
% and suppose that we have proved the existence of the lift $\tpsi_m\colon \ZcC{m}\to H_m(\cms)$
% with the aforementioned property.

\begin{defn}
\label{defn:variationsCm}
We introduce several variations of the notion of configuration space; see Figure \ref{fig:defcmsD}.
\begin{itemize} 
 \item The space $C_{1,m}(\S)$ is the subspace of $\mrS\times \cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots,P_m}}$ with $\bar P\neq P_i$ for all $i$; in other words it is
 the space of configurations of $m+1$ points, one of which is \emph{white} (meaning that
 it is distinguishable from the other points), whereas the other are \emph{black} and not distinguishable
 from each other.
%  \item The space $\fmsD$ is the subspace of $\mrS\times\cmsD$ containing all configurations
%  $(q;\set{p_1,\dots,p_m})$ where either $q\in\D$ may coincide with some $p_i$ (which can appear
%  with multipliciy $\geq 1$), or $q\not\in\D$ must be distinct from all $p_i$'s.
%  Again $q$ is thewhite point
 \item The space $\fmstwoD$ is the subspace of $\mrS\times\cms$ containing all configurations
 $\pa{\bar P;\set{P_1,\dots, P_m}}$ where either $\bar P\in\D$ may coincide with \emph{exactly}
 one $P_i$, or
 $\bar P\not\in \D$ must be distinct from all $P_i$'s.
 Again $\bar P$ is called the white point.
 %We see that $C_{1,m}\subset\fmstwoD$ is an open subspace.
 \item The space $\cmstwoD$ is the subspace of $\SP^{m+1}(\mrS)$ of configurations where either all $m+1$ points
 are distinct, or there is exactly one point \emph{inside $\D$} with multiplicity $2$ and $m-1$ other points,
 somewhere in $\mrS$, with multiplicity $1$.
\end{itemize}
 \end{defn}

\begin{figure}\centering
 \includegraphics{figures/defcmsD.pdf}
 \caption{A configuration in each of the space introduced in Definitions \ref{defn:cmsD} and \ref{defn:variationsCm}.
 Whenever a multiplicity is not specified, it is equal to 1.}
\label{fig:defcmsD}
\end{figure}

We have the following inclusions:
\[
C_{1,m}(\S)\subset\fmstwoD\subset\mrS\times\cmsD\subset\mrS\times\SP^m(\mrS);
\]
\[
C_{m+1}(\S)\subset\cmstwoD\subset C_{m+1}^{\D}(\S)\subset\SP^{m+1}(\mrS).
\]
All these spaces are
manifolds of dimension $2m+2$ and all inclusions are open.

@@

We can multiply both
cycles and the homology between them by the cycle $[c_1]$: the result are the two homologous cycles
$[c_1]\cdot \fc$ and $[c_1]\cdot\ldots\cdot[c_{m+1}]$ in $C_{m+1}^{\D}(\S)$: both cycles and the
homology between them are supported on $\N$. Note now that the
cycle $[c_1]\cdot\fc$ lives in $\cmstwoD$, so the first lift is done and we can now
deal with the second lift.

There is a natural map $\p\colon \fmstwoD\to \cmstwoD$, which converts the white point
into a black point. This map restricts to %maps $\fmstwoD\to\cmstwoD$ and 
a map $C_{1,m}(\S)\to C_{m+1}(\S)$, so that
% the latter being even a $m+1$-fold covering map;
we have a commutative diagram
 \begin{equation}\label{eq:cmstwodiagram}
  \begin{CD}
   C_{1,m}(\S) @>\subset >> \fmstwoD %@>\subset >> \fmsD
\\   @V\p VV @V\p VV %@V\p VV
\\   C_{m+1}(\S) @>\subset >> \cmstwoD %@>\subset >> C_{m+1}^D(\S)
   \end{CD}
\end{equation}

\begin{defn}
\label{defn:falsediagonals} 
Let 
\[
\Dmone=\cmstwoD\setminus C_{m+1}(\S)
\]
and similarly
\[
\Donem=\fmstwoD\setminus C_{1,m}(\S).
\]
We note that $\p$ restricts to a homeomorphism $\Donem\to\Dmone$. Moreover both $\Dmone\subset\cmstwoD$
and $\Donem\subset\fmstwoD$ are closed submanifolds of codimension $2$, and the map $\p$ restricts to
a $2$-fold ramified covering between their respective normal bundles.
\end{defn}

Diagram \eqref{eq:cmstwodiagram} induces a commutative diagram in homology
\begin{equation}
 \label{eq:fivediagram}
\minCDarrowwidth15pt
 \begin{CD}
  @. H_{m+1}\pa{\fmstwoD} @>>> H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)}\\
  @. @V\p_*VV @V\p_*VV\\
  H_{m+1}\pa{ C_{m+1}(\S)} @>>> H_{m+1}\pa{\cmstwoD} @>>> H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}
 \end{CD}
\end{equation}

@@

We first note that there is a lift of $[c_1]\cdot\fc$ to a cycle $[c_1]\otimes\fc$ in $\pa{\fmstwoD}$:
this is defined by declaring the point in $[c_1]\cdot\fc$ that spins around $c_1$
to be white. We then note that the right vertical map
\[
\p_*\colon H_{m+1}\pa{\fmstwoD, C_{1,m}(\S)} \to H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}
\]
can be rewritten, after using excision to tubular neighborhoods of $\Donem$ and $\Dmone$ respectively,
and the Thom isomorphism, as a map
\[
 H_{m-1}(\Donem)\to H_{m-1}(\Dmone).
\]
The latter map is multiplication by $2$, after identifying $\Dmone$ and $\Donem$ along $p$:
indeed the normal bundle of $\Donem$ is a
double covering of the normal bundle of $\Dmone$, hence the Thom class of the first disc
bundle corresponds to twice the Thom class of the second disc bundle. We are working
with coefficients in $\Z_2$, so multiplication by $2$ is the zero map.

Therefore the image of the cycle $[c_1]\otimes \fc$ along the diagonal of the square
in diagram \eqref{eq:fivediagram} is zero; hence the image of $[c_1]\cdot\fc$ in $H_{m+1}\pa{\cmstwoD,C_{m+1}(\S)}$
is zero; hence the homology class of $[c_1]\cdot\fc$ comes from $H_{m+1}(C_{m+1}(\S))$. More
precisely, there exists a cycle $\fc'$ in $C_{m+1}(\S)$ such that $\pa{\iota_{m+1}}_*(\fc')$ is homologous
to $[c_1]\cdot\fc$.

To prove Lemma \ref{lem:tpsiwithproperties} we need to find a good cycle and
a good homology, namely two that are supported on $\N$: a priori both $\fc'$ and the homology
between $\pa{\iota_{m+1}}_*(\fc')$ and $[c_1]\cdot\fc$ are only supported on $\mrS$.

This can be done by replacing, in the whole argument of the proof, the surface $\mrS$ with the surface $\N$.
We can define configuration spaces as in Definition \ref{defn:variationsCm} also for the open surface
$\N$, and we can repeat the argument considering $\N$ as the \emph{ambient surface}:
indeed we only needed a surface containing $\D$ and all curves $c_1,\dots,c_{m+1}$.
